---
title: "Visit Scotland Final Project"
output: html_notebook
author: "Louise Shambrook"
---

### Loading required libraries
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

### Notes on data cleaning
I will be removing the feature_code from all datasets, as this is the country code for Scotland. I will also be removing the column measurement, as this is irrelevant and duplicated information. 

On all datasets, there is an "All" response. The data are survey responses, and therefore it is reasonable to assume that "all" was an actual response. Therefore, I will not drop it / change it to NA, but it also does not give any insight to my analysis and I will be removing it from my plots etc.

For most of the datasets, I have simplified the units and reordered the columns, so the datasets are easier to interpret.

For the activities dataset, I have simplified the responses, and dropped the original column.

For the transport dataset, I have grouped the types, and dropped the original column.

For the regional_domestic dataset, I have split the unit column into the nights/visits, and pounds, as the nights/visits are measured in thousands, and the pounds are measured in millions.

For the location dataset, I have also simplified the responses, and dropped the original column. 

### Loading cleaned data
```{r}
# data has been cleaned in clean_data script

activities <- read_csv("clean_data/activities_clean.csv")
accomodation <- read_csv("clean_data/accomodation_clean.csv")
demographics <- read_csv("clean_data/demographics_clean.csv")
location <- read_csv("clean_data/location_clean.csv")
regional_domestic <- read_csv("clean_data/regional_domestic_clean.csv")
transport <- read_csv("clean_data/transport_clean.csv")
```

### Business questions to explore

- What kind of tourism activity generates the highest income for Scotland?
- Which kind of activity generates the most visits to sites?
- Is there a particular key demographic of visitors we should be focusing on? Who spends the most? Who visits the most?
- What type of locations receive the most visits? What kinds of locations do people spend the most in?
- Is there a particular method of travel our visitors arrive by? Some of our locations are remote and not easily accessible by public transport, so this would be good information to have.
- When looking at regional tourism, what insights can we gain? Where are our visitors from? How do they differ in the money they spend, nights they spend in Scotland, or number of visits they do?
- What type of accomodation is most popular? Is there a difference by location?
- How have general tourism rates changed over time?
- Is there any way to predict spending and visits for Visit Scotland?

### Initial exploratory data analysis

```{r}
# checking for NA values (aside from the "all" values)
summary(accomodation)
summary(activities)
summary(demographics)
summary(domestic_tourism)
summary(location)
summary(transport)


```

```{r}
# Q1. What kind of tourism activity generates the highest income for Scotland?
activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  arrange(desc(highest_income))

```

```{r}
# Q2. Which kind of activity generates the most visits to sites?
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  arrange(desc(highest_income))
```

```{r}
# Q3. 
# Is there a particular key demographic of visitors we should be focusing on?
# as there is no inter-relating data (when you isolate one variable, there is no
# data for the others), I will have to analyse each variable individually.

demographics %>%
  filter(age != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(age) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

demographics %>%
  filter(age != "All" & breakdown_of_domestic_tourism != "Visits") %>%
  group_by(age) %>%
  summarise(most_spend = sum(value)) %>%
  arrange(desc(most_spend))

demographics %>%
  filter(marital_status != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(marital_status) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# This doesn't seem very relevant; it's a binary definer. Especially with the younger generation, this may not be very important?

demographics %>%
  filter(gender != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(gender) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# This is pretty much the same as with married. They're pretty even, and I'm not sure they'll matter too much.

demographics %>%
  filter(employment_status != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(employment_status) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# Most people are employed, perhaps not unexpected, as various levels of employment were not available.

demographics %>%
  filter(children != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(children) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# Most people report no to children.

demographics %>%
  filter(access_to_car != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(access_to_car) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# This is an interesting and useful one. Further down, there is a question about travel, and here by far the majority report having access to a car.

demographics %>%
  filter(social_grade != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(social_grade) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# This mostly speaks for itself; the biggest caveat to this is that there was no data dictionary with this dataset, so I can only assume that it goes top down, but this is not known.

```


```{r}
# Q 3.2
# Who spends the most?
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & age != "All") %>%
  group_by(age) %>%
  summarise(highest_spend = sum(value)) %>%
  arrange(desc(highest_spend))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & marital_status != "All") %>%
  group_by(marital_status) %>%
  summarise(highest_spend = sum(value)) %>%
  arrange(desc(highest_spend))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & gender != "All") %>%
  group_by(gender) %>%
  summarise(highest_spend = sum(value)) %>%
  arrange(desc(highest_spend))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & employment_status != "All") %>%
  group_by(employment_status) %>%
  summarise(highest_spend = sum(value)) %>%
  arrange(desc(highest_spend))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & children != "All") %>%
  group_by(children) %>%
  summarise(highest_spend = sum(value)) %>%
  arrange(desc(highest_spend))

# The direct answer is people who identify as married - but there are some caveats to this answer. It is only possible to explore one variable at a time, and therefore, the identity of 'married' should be questioned. Additionally, because of the possible survey responses, marital_status only has two choices, whereas age is divided into decades starting at 16. So to best compare, age should be split/merged into two categories, or age is really the only identifier, as it has a range of reasonable options. 
```


```{r}
# Q 3.3
# Who visits the most?
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & age != "All") %>%
  group_by(age) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & marital_status != "All") %>%
  group_by(marital_status) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & gender != "All") %>%
  group_by(gender) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & employment_status != "All") %>%
  group_by(employment_status) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & children != "All") %>%
  group_by(children) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))

# Answer here is "no- without children.". But there is the same caveat as with the question above. 
```

```{r}
# Q4
# What type of locations receive the most visits?
location %>%
  filter(type_of_location_visited != "Other") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_visits = sum(value)) %>%
  arrange(desc(total_visits))

```

```{r}
# Q4.2
# What kinds of locations do people spend the most in?
location %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_spend = sum(value)) %>%
  arrange(desc(total_spend))
```

```{r}
# Q5
# Is there a particular method of travel our visitors arrive by?
transport %>%
  filter(breakdown_of_domestic_tourism == "Visits" & travel_type != "Other") %>%
  group_by(travel_type) %>%
  summarise(total_visits = sum(value)) %>%
  arrange(desc(total_visits))

# According to my analysis, most people use a car. This is especially underpinned by the demographics, where xx people report having access to a car.

# Note: the question / answer / analysis here from above, from demographics, with access_to_car, would be good to cover under transport.

```

```{r}
#Q6
# When looking at regional tourism, what insights can we gain? 
# How do they differ in
# the money they spend
# nights they spend in Scotland, or
# number of visits they do?

regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(region_of_residence) %>%
  summarise(total_spend = sum(value))

regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Nights") %>%
  group_by(region_of_residence) %>%
  summarise(total_units = sum(value))
  
regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  group_by(region_of_residence) %>%
  summarise(total_units = sum(value))

# by spend, england is the biggest.
# by nights, england is also the biggest - but the difference between england and scotland is less.
# by visits, scotland is bigger than england. This indicates local visiting.
```

```{r}
# Q6.1
# Where are our visitors from?
regional_domestic %>%
  group_by(region_of_residence) %>%
  summarise(breakdown_of_visit_countries = sum(value)) %>%
  arrange(desc(breakdown_of_visit_countries))

# Most visitors come from all over GB, i.e. including Wales. (it doesn't state the UK, but GB - but in general terms GB and UK are used interchangably - however, without a data dictionary, I'll have to go by the wording).

regional_domestic %>%
  group_by(region_of_residence, breakdown_of_domestic_tourism) %>%
  summarise(breakdown_of_visit_countries = sum(value))

# This provides a nicer breakdown.
```

```{r}
# Q7.
# What type of accomodation is most popular?
accomodation %>%
  group_by(accommodation_type_and_occupancy) %>%
  summarise(popular_accomodation = sum(value)) %>%
  arrange(desc(popular_accomodation))

# Taking the direct result, is hotels. But I think there are duplicates, somehow so it may need cleaning, and additionally, this is reported in percentage but I've only summed it so I can't report it like this - I'll need to do some further cleaning / wrangling.

# Note: these are reported in percentage; this is important to note on the plots and when talking about the figures

# Note: this is like the demographics; where there is data in one variables, the others have NA, i.e. 'All'.
```

```{r}
#Q7.1
# Is there a difference by location?
accomodation %>%
  filter(location != "Other") %>%
  group_by(location, accommodation_type) %>%
  summarise(popular_accomodation = sum(value)) %>%
  arrange(desc(popular_accomodation))

# So, people are staying in cities.

```

```{r}
# Q8.
# How have general tourism rates changed over time?
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

location %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

# There's a trend of visits increasing in 2017, and declining again.

activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

location %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = year, y = value)) +
  geom_col()

# Spend is more fluctuating. 

```


### Data visualisations for presentation

Transport
```{r}
# "Is there a particular method of travel our visitors arrive by?".

transport %>%
  filter(breakdown_of_domestic_tourism == "Visits" & travel_type != "Other") %>%
  group_by(travel_type) %>%
  summarise(total_visits = sum(value)) %>%
  ggplot(aes(x = reorder(travel_type, -total_visits), y = total_visits, fill = "blue")) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "Type of transport",
       y = "Total visits (million)",
       title = "Total visits by transport type 2013-2019") +
  scale_fill_manual(values = "#6B085E")

```
These are the sum of million visits in 2019, where visitors reported travelling by X.


```{r}
demographics %>%
  filter(access_to_car != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(access_to_car) %>%
  summarise(total_car = sum(value)) %>%
  ggplot(aes(x = access_to_car, y = total_car, fill = "blue")) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Year",
       y = "Total visits (million)",
       title = "Visits by demographic (access to car) 2013-2019",
       fill = "Access to car") +
  scale_fill_manual(values = "#6B085E")

```
This is underpinned by the finding that across the the (unit) million visits between 2013 and 2019, the majority of visitors report having access to a car.


Locations
```{r}
# "what type of locations receive the most visits?"

location %>%
  filter(breakdown_of_domestic_tourism == "Visits"
         & type_of_location_visited != "Other") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_visits = sum(value)) %>%
  ggplot(aes(x = type_of_location_visited, y = total_visits, fill = "blue")) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Type of location",
       y = "Total visits (million)",
       title = "Total visits by type of location 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
The sum of million visits in 2019 by location. 

```{r}
# "what kinds of locations do people spend the most in?"
location %>%
  filter(breakdown_of_domestic_tourism == "Expenditure"
         & type_of_location_visited != "Other") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_visits = sum(value)) %>%
  ggplot(aes(x = type_of_location_visited, y = total_visits, fill = "blue")) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Type of location",
       y = "Total spend (million £)",
       title = "Total spend by type of location 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
Sum of spend in 2019 in millions, by location.

```{r}
accomodation %>%
  filter(location != "Other") %>%
  group_by(location, accommodation_type) %>%
  summarise(popular_accomodation = sum(value)/100) %>%
  arrange(desc(popular_accomodation)) %>%
  ggplot(aes(x = location, y = popular_accomodation, fill = "blue")) +
  geom_col() +
  theme(legend.position = "none") +
  labs(x = "Type of location",
       y = "Percentage",
       title = "Average occupancy in accommodation by location type 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
This lines up with accommodation choices; in 2019, most people chose to visit urban areas.
(note to self: this also answers 'Is there a difference (in accommodation) by location?')


Accommodation
```{r}
# "What type of accomodation is most popular?"
accomodation %>%
  group_by(accommodation_type, occupancy_type) %>%
  summarise(popular_accomodation = sum(value)/100) %>%
  ggplot(aes(x = accommodation_type, y = popular_accomodation, fill = occupancy_type)) +
  geom_col(position = "dodge", color = "black") +
  labs(x = "Type of accommodation",
       y = "Percentage",
       title = "Average occupancy by accommodation type 2013-2019",
       fill = "Occupancy type") +
  scale_fill_manual(values = c("Room Occupancy" = "#6B085E",
                               "Unit Occupancy" = "#6B085E",
                               "Bed Occupancy" = "#905E88",
                               "Pitch Occupancy" = "#6B085E"))
  
```
2012-2019; hotel is by far the highest.

2019 data, from location set. When breaking down by location (as above), an interesting finding is that in the urban area, B&B's are just as popular as hotels. Maybe down to a certain B&B site?


```{r}
# trying to figure out the correct way of displaying the accommodation data
accomodation %>%
  filter(location != "Other") %>%
  group_by(location) %>%
  summarise(popular_accomodation = sum(value)/100) %>%
  ggplot(aes(x = location, y = popular_accomodation, fill = "blue")) +
  geom_col(position = "dodge") +
  theme(legend.position = "none") +
  labs(x = "Location",
       y = "Percentage",
       title = "Average occupancy by location type 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
Breaking the accommodation down by area, the most popular location type is urban areas.



```{r}
accomodation %>%
  filter(location == "Urban area" & accommodation_type == "Guest House/B&B") %>%
  group_by(year) %>%
  summarise(popular_accomodation = sum(value)/100) %>%
  ggplot(aes(x = year, y = popular_accomodation)) +
  geom_line() +
  labs(x = "Year",
       y = "Percentage",
       title = "Average occupancy by Urban Area and Guest House/B&B")
```
As B&B's are the second-most popular type, and urban areas are most popular - I wanted to see the trend over time. It is of significant interest that these are steadily increasing. 


Activities
```{r}
# What kind of tourism activity generates the highest income?
activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & activity_type != "Other") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  ggplot(aes(x = reorder(activity_type, -highest_income), y = highest_income, fill = "blue")) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "Type of activity",
       y = "Sum of spend (million £)",
       title = "Total spend by activity type 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
2013-2019, sum of value in million pounds. Highest is eating out (with shopping and visiting family/friends following)

```{r}
# Which kind of activity generates the most visits?
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits" & activity_type != "Other") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  ggplot(aes(x = reorder(activity_type, -highest_income), y = highest_income, fill = "blue")) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "Type of activity",
       y = "Visits (million)",
       title = "Total visits by activity type 2013-2019") +
  scale_fill_manual(values = "#6B085E")

```
2013-2019, sum of million visits. Highest is visits with family/friends. Eating out coming in second, and outdoor leisure in third.

Demographics 
```{r}
# Is there a particular key demographic of visitors we should be focusing on?

demographics %>%
  filter(age != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(age) %>%
  summarise(most_visits = sum(value)) %>%
  ggplot(aes(x = age, y = most_visits, fill = "blue")) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "Age group",
       y = "Visits (million)",
       title = "Total vists by age group 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
Yes, and no. The data from this dataset isn't great for analysis, because a lot of inferences have to be drawn. A key insight to be drawn from this is to complete further research / get better data.

With that said - age is a good variable and gives good insights. Across pretty much everything, the reflection from the data is that the younger generations are the ones attending and spending money.


```{r}
# breaking the age group by year
demographics %>%
  filter(age == "16-24 years" & breakdown_of_domestic_tourism == "Visits") %>%
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  labs(x = "Year",
       y = "Visits (million)",
       title = "Vists by age 16-24 across 2013 - 2019")
```


```{r}
demographics %>%
  filter(age == "55-64 years" & breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  labs(x = "Year",
       y = "Spend (million £)",
       title = "Spend by age group 55-64 across 2013 - 2019") 

demographics %>%
  filter(age == "65 years and over" & breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = year, y = value)) +
  geom_line() +
  labs(x = "Year",
       y = "Visits (million)",
       title = "Spend by age group 65+ across 2013 - 2019")
```


```{r}
# Who spends the most?
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & marital_status != "All") %>%
  group_by(marital_status) %>%
  summarise(highest_spend = sum(value)) %>%
  ggplot(aes(x = marital_status, y = highest_spend, fill = "blue")) +
  geom_col(width = 0.4) +
  theme(legend.position = "none") +
  labs(x = "Is married?",
       y = "Spend (million £)",
       title = "Total spend 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
This is where the data is limited, because the survey / options for respondents was limited, so marital_status can't really be fairly compared. However, going by the data strictly (because this does still reflect something, and even when grouping age so that they could be compared, the "married" response comes out higher).


```{r}
# Who visits the most?
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & children != "All") %>%
  group_by(children) %>%
  summarise(most_visits = sum(value)) %>%
  ggplot(aes(x = children, y = most_visits, fill = "blue")) +
  geom_col(width = 0.4) +
  theme(legend.position = "none") +
  labs(x = "Has children?",
       y = "Visits (million)",
       title = "Total vists 2013-2019") +
  scale_fill_manual(values = "#6B085E")
```
This is the same issue, but also response / relevance, as the above.


Regional demographics

```{r}
# When looking at regional tourism, what insights can we gain? Where are our visitors from? How do they differ in the money they spend, nights they spend in Scotland, or number of visits they do?

regional_domestic %>%
  filter(breakdown_of_domestic_tourism != "Expenditure") %>%
  ggplot(aes(x = region_of_residence, y = value, fill = breakdown_of_domestic_tourism)) +
  geom_col(position = "dodge") +
  labs(x = "Region",
       y = "Visits/Nights (thousands)",
       title = "Nights and visits by region",
       fill = "Measure")+
  scale_fill_manual(values = c("Nights" = "#6B085E",
                               "Visits" = "#905E88"))

regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  ggplot(aes(x = region_of_residence, y = value, fill = breakdown_of_domestic_tourism)) +
  geom_col(position = "dodge") +
  theme(legend.position = "none") +
  labs(x = "Region",
       y = "Spend (million £)",
       title = "Spend by region") +
  scale_fill_manual(values = "#6B085E")
  
```


Change over time
```{r}
# How have general tourism rates changed over time?

transport %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  group_by(year) %>%
  summarise(total_year = sum(value)) %>%
  ggplot(aes(x = year, y = total_year)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(x = "Year",
       y = "Sum of visits (millions)",
       title = "Tourism rates (by transport, visits) 2013-2019")

demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(year) %>%
  summarise(total_year = sum(value)) %>%
  ggplot(aes(x = year, y = total_year)) +
  geom_line() +
  theme(legend.position = "none") +
  labs(x = "Year",
       y = "Sum of spend (million £)",
       title = "Tourism rates (by demographics, spend) 2013-2019")
```
The *general* picture over time, is that for visits, it increases until 2017, and then declines (some more than others).
For expenditure, it increases until 2017, declines in 2018, and increases again in 2019.

Predicting change
This has been plotted in python. The model was inaccurate. I also tried to clean/wrangle the data so that I could plot it in R, but this didn't work either.


### Summary and descriptive statistics for the presentation

Transport slides
```{r}
# total transport visits who use other types of transportation
transport %>%
  filter(breakdown_of_domestic_tourism == "Visits" & travel_type != "Other") %>%
  summarise(total_value = sum(value))

# total transport visits who use car
transport %>%
  filter(breakdown_of_domestic_tourism == "Visits" & travel_type == "Car") %>%
  summarise(total_value = sum(value))

```
605.3/930.9
65%


```{r}
# total from visits who report having access to a car
demographics %>%
  filter(access_to_car == "Access to car (1+)" & breakdown_of_domestic_tourism == "Visits") %>%
  summarise(total_value = sum(value))

# total from visits
demographics %>%
  filter(access_to_car != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  summarise(total_value = sum(value))

```
768/940
81.7% report having access to a car


Location slides
```{r}
# total who visit city/large town
location %>%
  filter(breakdown_of_domestic_tourism == "Visits" & type_of_location_visited == "City/large town") %>%
  summarise(total_value = sum(value))

# total who visit (by location)
location %>%
  filter(breakdown_of_domestic_tourism == "Visits" & type_of_location_visited != "Other") %>%
  summarise(total_value = sum(value))
```
431/1032
41.7%

```{r}
# average accommodation occupancy in urban areas
accomodation %>%
  filter(location == "Urban area") %>%
  summarise(popular_accomodation = sum(value)/100)

```
68%

Accommodation slides
```{r}
# average accommodation occupancy
accomodation %>%
  group_by(accommodation_type, occupancy_type) %>%
  summarise(popular_accomodation = sum(value)/100)

```
Hotels are the highest, with 66.27%

Accommodation slides
```{r}
# average accommodation occupancy in urban areas
accomodation %>%
  filter(location == "Urban area") %>%
  summarise(popular_accomodation = sum(value)/100)

```

Other point is urban B&B increasing.


Demographics slides
```{r}
# sum of expenditure by is_married
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & marital_status == "Married") %>%
  summarise(popular_accomodation = sum(value))
# 23928 (million)

# sum of expenditure 
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & marital_status != "All") %>%
  summarise(popular_accomodation = sum(value))
# 35751

```
23928/35751
Married - 66.9% spend

Demographics
```{r}
# sum of visits by have_children = no
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & children == "No - Children") %>%
  summarise(popular_accomodation = sum(value))
683

# sum visits
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & children != "All") %>%
  summarise(popular_accomodation = sum(value))
940

```
683/940 - 72%


Demographics slides - age
```{r}
# 16-24 year old - sum of visits
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & age == "16-24 years") %>%
  summarise(popular_accomodation = sum(value))

# sum of visits (all age groups)
demographics %>%
  filter(breakdown_of_domestic_tourism == "Visits" & age != "All") %>%
  summarise(popular_accomodation = sum(value))
```
184/946 - 19.4%


Demographics slides - Older age groups
```{r}
# sum of spend of 55-64 year olds
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & age == "55-64 years") %>%
  summarise(total = sum(value))
5792

# sum of spend 65+
demographics %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & age == "65 years and over") %>%
  summarise(total = sum(value))
5070
```
5070+5792 - 10862 total

```{r}
# finding the difference between 16-24 and older age groups
demographics %>%
  filter(age != "All" & breakdown_of_domestic_tourism == "Visits") %>%
  group_by(age) %>%
  summarise(most_visits = sum(value))
```
16-24 - 184
55-64 - 158
65+ - 164


Regional demographics
```{r}
# total visits where region is england
regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Visits" & region_of_residence == "England") %>%
  summarise(total = sum(value))

# total visits where region is scotland
regional_domestic %>%
  filter(breakdown_of_domestic_tourism == "Visits" & region_of_residence == "Scotland") %>%
  summarise(total = sum(value))

```
England - 109680
Scotland - 112132
Difference is 2452

Note; haven't used this number.

Activities
```{r}
# total spend of eating out activity
activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & activity_type == "Eating out") %>%
  summarise(total = sum(value))

# total spend of all activities
activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure" & activity_type != "Other") %>%
  summarise(total = sum(value))

```
13409/75277

Eating out accounts for 13409 million (17.8%)

```{r}
# sum of visits for 'visit family/friends' activity
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits" & activity_type == "Visit family/friends") %>%
  summarise(total = sum(value))

# sum of visits for all activities
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits" & activity_type != "Other") %>%
  summarise(total = sum(value))
```
315/1694

Visiting family/friends 315 million – 18.5%



### Notes on VisitScotland updating their data
Following the challenges with data modelling, I completed further research on VisitScotland's site. They also complete another piece of research, called the Great Britain Tourism Survey. They have noted (https://www.visitscotland.org/binaries/content/assets/dot-org/pdf/research-papers-2/uk-visitors/gbts-gbdvs-update-nov21.pdf) that from 2020 onwards, they will be changing the structure of the data. It seems that they have found the same issues as myself, and are putting similar changes to the ones I recommend in place.

However, there is no data available yet, as travel has been severely impacted due to Covid.

https://www.visitscotland.org/binaries/content/assets/dot-org/pdf/research-papers-2/uk-visitors/gbts-gbdvs-update-nov21.pdf

### Conclusions and summary of insights
Specifically, the insights from my analysis were that the majority of visitors travel by and have access to a car. Additionally, there is a trend of visitors visiting cities; this is evident by exploring both the location data, but also by exploring the accommodations visitors use. Continuing on this thread, B&B's and hotels are the most popular type of accommodation. Additionally, there is an interesting intersection in data between urban areas, and the average occupation of B&B's, as this has been increasing since 2016. The overall profile of a visitor to Scotland is someone who does not have children, and is married; however, a key caveat to this is that the data collection method in demographics is binary, and therefore there are limited options. A better way to profile visitors is by age; 16-24 is the most popular age group, but this is declining, and so for an age group to target, the age group of 55+ is most prolific, as the attendance of these have been steadily increasing since 2015. Data is also profiled by home country, and here England is the dominant feature; however, this is also a dataset where there are limited options, as Wales and Northern Ireland are not included. Lastly, the most common activities in Scotland by visitors are eating out, and visiting family and friends

### Main recommendations:
From my analysis and insights, I have recommended that VisitScotland can be confident that the majority of their customers travel by and have access to car, so any promotions that they may have planned which involve remote locations, they can confidently proceed with.
I recommended that VisitScotland develop partnerships with B&B providers, particularly in urban areas, which is an area that they should focus on. In terms of profiling a demographic, the age group of 55+ is the age group I recommend to target.
The key takeaway from my analysis however, is to restructure their data collection, as there were shortcomings in two datasets which made analysis a challenge. I also recommended that they collect data more frequently, e.g. monthly, as this would allow for predictive models to be built, which is not currently possible. 



