---
title: "Visit Scotland Final Project CodeClan 2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
  pdf_document: default
---
Main project draft output

```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tsibble)
library(lubridate)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(leaflet)
library(tsibbledata)



```

```{r}
READ IN DATA

```


# 1. Documentation

#### Domain knowledge (1.1)



  
  
#### Business requirements (1.2)






#### Business processes and data flow (1.3)




#### Data visualisation as a tool for decision-making (1.4)





#### Data types (1.5)




#### Data quality and data bias (1.6)




# 2. Data cleaning

#### Preparing data for visualisation (1.7)

Cleaning notes

I will be removing the feature_code from all datasets, as this is the 
country code for Scotland. I will also be removing measurement, as this is irrelevant
and duplicated information. 

On all datasets, there is an "All" response. The data are survey responses, and therefore it is reasonable to assume that "all" was an actual response. Therefore, I will not drop it / change it to NA, but it also does not give any insight to my analysis and I will be removing it from my plots etc.

On the activities dataset, I initially wanted to group these (maybe I still will) - but have settled on definitely simplifying them.

On the transport dataset, I have grouped the types of travel, as there is data that is missed.

Outline cleaning and wrangling code





# 3. Data visualisation
#### Introduction to dataset and data ethics



#### Process and design 




#### Visualisations

Note to self - pick 5-10 plots to put in here and describe what they show





### Conclusions

