---
title: "Visit Scotland Final Project CodeClan 2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
  pdf_document: default
---
Main project draft output

```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tsibble)
library(lubridate)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(leaflet)
library(tsibbledata)



```

```{r}
READ IN DATA

```


# 1. Documentation

#### Domain knowledge (1.1)



  
  
#### Business requirements (1.2)






#### Business processes and data flow (1.3)




#### Data visualisation as a tool for decision-making (1.4)





#### Data types (1.5)




#### Data quality and data bias (1.6)




# 2. Data cleaning

#### Preparing data for visualisation (1.7)





Outline cleaning and wrangling





# 3. Data visualisation
#### Introduction to dataset and data ethics



#### Process and design 




#### Visualisations

Note to self - pick 5-10 plots to put in here and describe what they show





### Conclusions

