---
title: "Visit Scotland Final Project"
output: html_notebook
author: "Louise Shambrook"
---

- note to self - structure this document/report similarly to the bikes project

libraries
```{r}
### loading in libraries
library(tidyverse)
library(janitor)
library(lubridate)

```

Exploring the data for each dataset
```{r}
head(domestic_tourism)

domestic_tourism %>%
  filter(Units == "million pounds (GBP)")

# checking for NA values

summary(accomodation)
summary(activities)
summary(demographics)
summary(domestic_tourism)
summary(location)
summary(transport)

head(activities)
head(location)
head(transport)
head(domestic_tourism)
head(accomodation)
```

### Reading in cleaned data

```{r}
activities <- read_csv("clean_data/activities_clean.csv")
accomodation <- read_csv("clean_data/accomodation_clean.csv")
demographics <- read_csv("clean_data/demographics_clean.csv")
location <- read_csv("clean_data/location_clean.csv")
regional_domestic <- read_csv("clean_data/regional_domestic_clean.csv")
transport <- read_csv("clean_data/transport_clean.csv")
```

### Initial analysis -----------------------------------------------------------

Q1.
What kind of tourism activity generates the highest income for Scotland?
```{r}
activities %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  arrange(desc(highest_income))
```
The answer from this analysis, is going out for a meal. 

Which kind of activity generates the most visits to sites?
```{r}
activities %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  group_by(activity_type) %>%
  summarise(highest_income = sum(value)) %>%
  arrange(desc(highest_income))
```

Q3. 
Is there a particular key demographic of visitors we should be focusing on?
```{r}
# as there is no inter-relating data (when you isolate one variable, there is no
# data for the others), I will have to analyse each variable individually.

# age

demographics %>%
  filter(age != "All") %>%
  group_by(age) %>%
  summarise(most_visits = sum(value)) %>%
  arrange(desc(most_visits))
```


Q3.1 Who spends the most?
```{r}

```

Q3.2 Who visits the most?
```{r}

```


Q4.
What type of locations receive the most visits? 
```{r}
location %>%
  filter(breakdown_of_domestic_tourism == "Visits") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_visits = sum(value)) %>%
  arrange(desc(total_visits))
```

Q4.1
What kinds of locations do people spend the most in?
```{r}
location %>%
  filter(breakdown_of_domestic_tourism == "Expenditure") %>%
  group_by(type_of_location_visited) %>%
  summarise(total_spend = sum(value)) %>%
  arrange(desc(total_spend))
```

Q5.1
Is there a particular method of travel our visitors arrive by? Some of our locations are remote and not easily accessible by public transport, so this would be good information to have.
```{r}


how to count????
  
  
```

Q6.
When looking at regional tourism, what insights can we gain? How do they differ in the money they spend, nights they spend in Scotland, or number of visits they do?

```{r}


```

Q6.1
Where are our visitors from? 
```{r}

```

Q7.
What type of accomodation is most popular? Is there a difference by location?
```{r}

```

Q8.
How have general tourism rates changed over time?
```{r}

```




## Steps / questions / Plan

- Whilst answering each question, keep an eye out for details / small points - the overall story - remember the bikes project!
- Figure out how to predict spending for visit scotland
- Figure out how to predict visits for visit scotland
    - For these, what are my options? (linear, logistic, decision tree, forest, any others? Maybe clean and prepare the data in R, and then run the model in Python???)


### Main questions to answer --------------------------------------------------



Is there any way to predict spending and visits for Visit Scotland?


### Sub-questions to answer / points to explore, as I go along: