---
title: "Visit Scotland Final Project"
output: html_notebook
author: "Louise Shambrook"
---

libraries
```{r}
### loading in libraries
library(tidyverse)
library(janitor)
library(lubridate)

```

reading in data
```{r}
domestic_tourism <- read_csv("raw_data/regional_domestic_tourism.csv")
accomodation <- read_csv("raw_data/scottish_accomodation_occupancy.csv")
activities <- read_csv("raw_data/tourism_day_visits_activities.csv")
demographics <- read_csv("raw_data/tourism_day_visits_demographics.csv")
location <- read_csv("raw_data/tourism_day_visits_location.csv")
transport <- read_csv("raw_data/tourism_day_visits_transport.csv")

```

EDA
```{r}
head(domestic_tourism)

domestic_tourism %>%
  filter(Units == "million pounds (GBP)")
```

EDA on activities
```{r}
activities 
```
Notes on cleaning of activities
- janitor on names
- drop 'featurecode'
- drop measurement
- figure out how to group / sort this (the question is which type generates the most visits?)


EDA on demographic
```{r}
demographics %>%
  select(Age)


```
Cleaning on demographics
- janitor
- drop featurecode
- drop measurement
- there are many rows with "All" - seems to be an NA / other value. Need to decide what I'll do about that. 
- One of the questions is "who spends the most" (and visit) - spend and visit aren't features in that dataset. So I will have to think about joining, and how I do that.


EDA on accomodation
```{r}
accomodation

accomodation %>%
  distinct(`Weekday/Weekend`)
```

Notes on cleaning accomodation
- use janitor
- drop 'FeatureCode'
- drop 'measurement'


The plan is to eventually have all cleaning in one script, but for now, do by cleaning and question.


## Steps / questions / Plan

- Complete initial EDA; find out how much data cleaning needs to be completed

- Begin to complete wrangling so that each question can be answered
- Whilst answering each question, keep an eye out for details / small points - the overall story - remember the bikes project!
- Figure out how to predict spending for visit scotland
- Figure out how to predict visits for visit scotland
    - For these, what are my options? (linear, logistic, decision tree, forest, any others? Maybe clean and prepare the data in R, and then run the model in Python???)


### Main questions to answer

What kind of tourism activity generates the highest income for Scotland?

Which kind of activity generates the most visits to sites?

Is there a particular key demographic of visitors we should be focusing on? Who spends the most? Who visits the most?

What type of locations receive the most visits? What kinds of locations do people spend the most in?

Is there a particular method of travel our visitors arrive by? Some of our locations are remote and not easily accessible by public transport, so this would be good information to have.

When looking at regional tourism, what insights can we gain? Where are our visitors from? How do they differ in the money they spend, nights they spend in Scotland, or number of visits they do?

What type of accomodation is most popular? Is there a difference by location?

How have general tourism rates changed over time?

Is there any way to predict spending and visits for Visit Scotland?


### Sub-questions to answer / points to explore, as I go along: